\documentclass[12pt]{article}
\usepackage{amssymb,amsfonts,amsmath,mathtext,mathtools}
\usepackage{xfrac}
\usepackage{url, hyperref}
\usepackage[inline, shortlabels]{enumitem}


\newcommand{\avg}[1]{\langle{#1}\rangle}
\newcommand{\W}{\Omega}
\newcommand{\w}{\omega}
\newcommand{\nbar}{\bar n}
\newcommand{\rd}{\mathrm{d}}
\newcommand{\ddt}[1]{\frac{\rd{#1}}{\rd t}}
\newcommand{\D}{\Delta}
\newcommand{\geff}{\gamma_{eff}}

\title{Frequency Domain Method}

\begin{document}
	\maketitle
	\begin{abstract}
		A new method for searching for the electric dipole moment (EDM) of the deuteron and other nuclei is presented. When trying to measure the EDM in a storage ring environment, magnetic dipole moment (MDM) spin precession due to machine imperfections becomes the primary source of systematic error. The proposed method aims at providing a solution to the machine imperfection problem. The method is based on estimating the combined MDM + EDM spin precession angular velocity, in which the MDM contribution is due only to field imperfections. The MDM term is canceled in the final statistic by  adding angular velocity estimates from cycles with counter-circulating beams. Spin precession rate depends on the particle's effective Lorentz factor; the proposed method's core feature is a procedure for equalizing the effective Lorentz factors of the clockwise and counter-clockwise circulating beams, thus enabling the cancelation.
	\end{abstract}

\tableofcontents

\section{Introduction}

\section{Machine imperfection MDM spin precession}
\input{machine_imperfection_MDM_precession}

%\section{Spin Wheel}
%
%\subsection{Critique of the SW method}
%The Spin Wheel method is based on two equations:
%\begin{align}
%	\avg{\Omega_V}_x &= -\frac{Ze}{\gamma Am_pc}\left(\avg{B_x} - \frac{\avg{E_z}}{\beta}\right) \equiv 0, \label{eq:B-to-offset}
%	\shortintertext{and}
%	\avg{\Omega_x} &= -\frac{Ze}{Am_pc}\frac{1+a}{\gamma^2}\left(\avg{B_x} + \frac{\eta}{a}\avg{E_x}\right).\label{eq:B-to-s-prec}
%\end{align}
%Equation~\eqref{eq:B-to-offset} determines the relationship between the average radial magnetic field present in the beamline, and the resulting vertical shift of the beam's closed orbit:
%\begin{equation}
%	\avg{z} = \frac{1}{\avg{G_z}}\Big[\beta\avg{B_x} - \avg{E_z(0)}\Big].
%\end{equation}
%
%Equation~\eqref{eq:B-to-s-prec} connects the magnetic field to the beam's spin precession frequency.

\section{Frequency Domain Method}
\subsection{EDM estimator statistic}
\input{FD_EDM_statistic}

\subsection{Effective Lorentz factor}
%\input{ELF/ffective_lorentz_factor}
\input{ELF/effective_lorentz_factor_application}

\subsubsection{Calibration of the ELF}
\input{ELF/calibration}

\section{Statistical precision}
\input{statistical_precision}

\begin{thebibliography}{0}
%	\bibitem{Senichev:IPAC13}
%	Y. Senichev et al., ``Spin tune decoherence effects in Electro- and Magnetostatic Structures.''
%	Proceedings of IPAC 2013, Shanghai, China, pp. 2579-2581.
	\bibitem{Senichev:FDM}
	Y. Senichev, A. Aksentev, A. Ivanov, E. Valetov, ``Frequency domain method of the search for
	the deuteron electric dipole moment in a storage ring with imperfections,'' arxiv:1711.06512 [physics.acc-ph]
	\url{https://arxiv.org/abs/1711.06512}.
	\bibitem{Aksentev:Thesis}
	A. Aksentev, ``2D Frozen spin method of searching for the deuteron EDM in a storage ring.'' PhD thesis, NRNU ``MEPhI,'' Moscow, Russia.
	\url{http://collaborations.fz-juelich.de/ikp/jedi/public_files/theses/dissertation.pdf}
	\bibitem{Aksentev:IPAC19:SMP}
	A. Aksentev, Y. Senichev, ``Spin Motion Perturbation Effect on the EDM Statistic
	in the Frequency Domain Method,'' presented at the 10th International Particle Accelerator Conf. (IPAC'19),
	Melbourne, Australia, May. 2019.
	\url{https://ipac2019.vrws.de/papers/mopts011.pdf}
	\bibitem{Stats}
	A.E. Aksentev, Y.V. Senichev, ``Statistical precision in charged particle EDM search in storage rings.'' J Phys: Conf Ser. \textbf{941} (2017) 012083. 
	\url{https://iopscience.iop.org/article/10.1088/1742-6596/941/1/012083}
	
\end{thebibliography}

\end{document}