\documentclass{article}

\usepackage[widepage]{repsty}
\usepackage{subcaption}

\newcommand{\W}{\Omega}
\newcommand{\w}{\omega}
\newcommand{\MDM}{{}^\mathrm{MDM}}
\newcommand{\EDM}{{}^\mathrm{EDM}}
\newcommand{\CW}{{}^\mathrm{CW}}
\newcommand{\CCW}{{}^\mathrm{CCW}}



\begin{document}
\title{Imperfect structures}
\section{Analysis of systematic errors}
\subsection{E+B element tilt}
The E+B element tilt about the vertical axis is simulated as described in~\cite{Valetov:Thesis}, by adding spin rotations about the radial axis associated with the parasitic $B_X$-, and $E_y$-fields. The former is supposed a result of the tilt proper, while the latter is due to Lorentz force compensation, i.e.,
\begin{align}
  B_x &= B\tan\theta, \\
  E_y &= \frac{\chi_E}{\chi_m} B_x.
\end{align}

By doing it this way, we change only the spin dynamics, while maintaining the beam dynamics (i.e. the closed orbit) intact. 

The imperfection addition to the radial precession frequency is
\[
\Delta\Omega\MDM = -\frac{q}{m}\bkt*{G B_x + \bkt{\frac{1}{\beta^2} - 1 - G}\beta E_y}.
\]


\begin{thebibliography}{99}
\bibitem{BNL_proposal}
  D. Anastassopoulos, V. Anastassopoulos, D. Babusci. AGS Proposal: Search for a permanent electric dipole moment of the deuteron nucleus at the $10^{−29} ~ e\cdot cm$ level. [Internet]. BNL; 2008 [cited 2016 Nov 25]. Available from: \url{https://www.bnl.gov/edm/files/pdf/deuteron_proposal_080423_final.pdf}
\bibitem{Mane:SpinWheel}
  S. R. Mane. Spin Wheel. arXiv:150901167 [physics] [Internet]. 2015 Sep 3 [cited 2018 Sep 28]; Available from: http://arxiv.org/abs/1509.01167
\bibitem{Senichev:FDM}
  Frequency domain method of the search for the deuteron electric dipole moment in a storage ring with imperfections

\bibitem{Senichev:StorageRingMethod}
  Yurij Senichev. Search for the Charged Particle Electric Dipole Moments in Storage Rings. In: 25th Russian Particle Accelerator Conference (RuPAC’16), St Petersburg, Russia, November 21-25, 2016 [Internet]. JACOW, Geneva, Switzerland; 2017 [cited 2017 Apr 5]. p. 6–10. Available from: \url{http://accelconf.web.cern.ch/AccelConf/rupac2016/papers/mozmh03.pdf}

\bibitem{Senichev:Decoh}
  Senichev Y, Zyuzin D. SPIN TUNE DECOHERENCE EFFECTS IN ELECTRO- AND  MAGNETOSTATIC STRUCTURES. In: Beam Dynamics and Electromagnetic Fields [Internet]. Changhai, China: JACoW; 2013 [cited 2017 Jul 31]. p. 2579--2581. Available from: \url{https://accelconf.web.cern.ch/accelconf/IPAC2013/papers/wepea036.pdf}

\bibitem{Valetov:Thesis}
  Eremey Valetov. FIELD MODELING, SYMPLECTIC TRACKING, AND SPIN DECOHERENCE FOR EDM AND MUON G-2 LATTICES [Internet]. [Michigan, USA]: Michigan State University; Available from: \url{http://collaborations.fz-juelich.de/ikp/jedi/public_files/theses/valetovphd.pdf}


\end{thebibliography}
\end{document}
