\chapter*{Introduction}                         % Заголовок
\addcontentsline{toc}{chapter}{Introduction}    % Добавляем его в оглавление

\input{common/characteristic} % Характеристика работы по структуре во введении и в автореферате не отличается (ГОСТ Р 7.0.11, пункты 5.3.1 и 9.2.1), потому её загружаем из одного и того же внешнего файла, предварительно задав форму выделения некоторым параметрам

\textbf{Structure of the dissertation.} This dissertation consists of an introduction, three chapters,
a conclusion, and one appendix.

\textbf{Chapter one}: 
\begin{enumerate}[(1)]
	\item Introduces the Frozen Spin concept.
	\item Provides a classification of frozen spin (FS) storage ring (SR) EDM search methodologies.
	\item Classifies some problems encountered in any FS SR EDM experiment.
	\item Describes the 2D FS method which aims to provide a solution to those problems.
	\item Describes some lattices that could be used with the proposed method.
\end{enumerate}

In the \underline{\textbf{second chapter}} we alnalyze the problems outlined in the first chapter, and their solutions; simuilation results follow. 

Considered problems:
\begin{enumerate}[(1)]
	\item perturbations in the particle spin dynamics caused by betatron oscillations, and their effect on the EDM-statistic of the 2D FS method;
	\item spin decoherence in the frozen spin regime;
	\item properties ang magnitude of the EDM-faking spin precession systematic error induced by machine imperfections;
	\item the guide field flipping procedure used for the elimination of the systematic error within the 2D FS methodology.
\end{enumerate}

A section is dedicated to the investigation of the question of interpretation of the notion of the \emph{effective Lorentz factor} ($\geff$) introduced in the first chapter. 

This notion is the foundation of a great deal of the 2D FS methodology. It can be defined as follows: if two particles' $\geff$ are equal, then their spin dynamics are equivalent (specifically, their spin precession angular velocity vectors have equal orientations and magnitudes), regardless of the particulars of their orbital motion. 

Specifically, what enables us to exclude the EDM-faking MDM precession from the 2D FS EDM-statistic is the fixing of the $\geff$ characterizing the beam.

In \underline{\textbf{chapter three}} we highlighted some of the more important (in the context of the present research) technologies developed within the EDM search project carried out at the Cooler Synchrotron (COSY); we also described the results of the spin coherence time (SCT) optimization studies done at COSY during the April-May 2019 beamtime.

One phenomenon worth noting is the SCT change observed during prolonged (destructive) polarimetry measurements, which is (presumably) caused by the transition from the outer (halo) to the inner (core) beam layers. The observation of this phenomenon can be explained within the bounds of the sextupole spin decoherence suppression theory outlined in this dissertation.

In the \underline{\textbf{conclusion}}, we summarize some fo the main thesis results:
\input{common/concl}

The main body of the dissertation does not include our statistical analysis of the experiment; it is placed in \textbf{appendix} A. Two aspects of that analysis worth mentioning are: investigation of the possibility to use a non-uniform polarization sampling scheme in order to optimize the beam lifetime; determiation of the maximum effective measurement cycle duration. 

As a result, we came to the conclusion that the non-uniform sampling scheme is not practical, due to the nature of polarimetry measurement. Concerning the optimal measurement cycle length, it cannot exceed three beam polarization life times.
